Describing the scale and composition of calls for service in Detroit

### 2021-08-28 paper.Rmd
writing

A study from @ratcliffe2021policing in _Crime Science_ sought to describe the complexity and diversity of public demand for police services in Philadelphia, United States. The study was motivated by the recognition that police spend a considerable proportion of deployed resource resolving incidents which fall outside the traditional crime-fighting role of the police. Instead, the origins of public demand for the police can often be attributed to a lack of supply and/or accessibility failure in other organisations, such as those providing support for people requiring (mental) health assistance [@wood2021can; @van2017law]. The study's findings provided data-driven evidence to suggest that the police indeed spend considerable amounts of time dealing with incidents that could be resolved by other -- often more appropriate -- agencies. Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can]. In advance of recent contributions, this short paper aims to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).

@@

Each incident has a _calldescription_ variable which describes the nature of the call. Call descriptions involving administrative duties (e.g., "start of shift information"), completely unknown problems, and non-deployment (e.g., "employee call in / time off") were removed. This left 207 unique call description categories. In the interests of parsimony and ease of interpretation these categories were recoded into 101 broader call descriptions. ^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (xxxxxxxxxxxxxxxxxxxxxxxxxx). Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.] Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing. 2.8% of incidents were deemed classifiable. For each of the six demand types, an 'other' classification was generated for those incident types which consumed less than 0.2% of police time during the year. This left a total of 48 categories for the reported breakdowns, capturing 267 thousand individual public calls for service during 2019.

In the spirit of @ratcliffe2021policing, the frequency counts and proportional breakdowns of demand type classifications are reported in Table 1. This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of issues. In Detroit, just 46% of police time was spent dealing with calls for service relating to crime (55% in Philadelphia). The remainder of time was spent dealing with community issues (6%), public and mental health (7%), proactive activity (7%), quality of life incidents (19%) and traffic duties (13%). The proportional breakdown of time consumed for each incident, grouped by demand type classification, are visualised in Figure 1. Here, it is clear that a small number incident types tend to dominate each demand classification. While incident descriptions are not always comparable, this visualisation confirms the diversity and complexity of public demand for police services reported in @ratcliffe2021policing.

@@

Each incident also has corresponding latitude and longitude coordinates and a call timestamp. Due to irregularities in geocoding, likely due unknown locations for some calls, 2% of incidents were excluded from the spatial descriptive analyses.

@@

>>>>

A study from @ratcliffe2021policing in _Crime Science_ sought to describe the complexity and diversity of public demand for police services in Philadelphia, United States. The study was motivated by the recognition that police spend a considerable proportion of deployed resource resolving incidents which fall outside the traditional crime-fighting role of the police. Instead, the origins of public demand for the police can often be attributed to a lack of supply and/or accessibility failure in other organisations, such as those providing support for people requiring (mental) health assistance [@wood2021can; @van2017law]. The study's findings provided data-driven evidence to suggest that the police indeed spend considerable amounts of time dealing with incidents that could be resolved by other -- often more appropriate -- agencies. Comparable findings were subsequently reported in the United States across multiple (anonymised) police jurisdictions [@lum2021can]. In advance of recent contributions, this short paper uses comparable descriptive statistics to describe the public demand for police services in Detroit, United States, using open data and open code which can be used to reproduce the findings (see xxxxxxxxxxxxxxxxxxxxxxxxxxxx).

@@

Each incident has a _calldescription_ variable which describes the nature of the call. Call descriptions involving administrative duties (e.g., "start of shift information"), completely unknown problems, and non-deployment (e.g., "employee call in / time off") were removed. This left 207 unique call description categories. In the interests of parsimony and ease of interpretation these categories were recoded into 101 broader call descriptions. ^[We refer readers to the corresponding GitHub repository for further details on how call descriptions were combined (xxxxxxxxxxxxxxxxxxxxxxxxxx). Readers unfamiliar with R code can review the reference table entitled 'categorisation summary'.] Each call description was then categorised into the six demand type classifications used in @ratcliffe2021policing. 2.8% of incidents were deemed unclassifiable. For each of the six demand types, an 'other' classification was generated for those incident types which consumed less than 0.2% of police time during the year. This left a total of 48 categories for the reported breakdowns, capturing 267 thousand individual public calls for service during 2019.

In the spirit of @ratcliffe2021policing, the frequency counts and proportional breakdowns for each demand type classifications are reported in Table 1. This tells a comparable story to Philadelphia: a considerable proportion of citizen calls for service, and in turn, time spent by officers, is committed to a diverse array of issues. In Detroit, just 46% of police time was spent dealing with calls for service relating to crime (55% in Philadelphia). The remainder of time was spent dealing with community issues (6%), public and mental health (7%), proactive activity (7%), quality of life incidents (19%) and traffic duties (13%).

The proportional breakdown of time consumed for each incident, grouped by demand type classification, are visualised in Figure 1. Here, it is clear that a small number incident types tend to dominate each demand classification. While incident descriptions are not always comparable, this visualisation confirms the diversity and complexity of public demand for police services reported in @ratcliffe2021policing.

@@

@@

Each incident also has corresponding latitude and longitude coordinates to describe the incident location. 2% of incidents were excluded from the spatial descriptive analyses due to incomplete or unknown geographic information. Individual points were aggregated to 1000ft$^2$ grid cells laid over the city to generate basic descriptive heatmaps of counts during the year. These are visualised for each demand classification in Figure 3.
